### 2011-12-25 paper.tex
Convert instances of apropos(1) to apropos and change some active voice to passive voice. More work coming.

\documentclass[titlepage, a4paper, 12pt]{article}
\usepackage{listings}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[colorlinks=false]{hyperref}
\lstset{ %
language=sh,
basicstyle=\footnotesize,
numbers=none,
backgroundcolor=\color{white},
showspaces=false,
showstringspaces=false,
showtabs=false,
frame=single,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
title=\lstname
}
\title{Apropos Replacement \newline
(Extended  Abstract)}
\author{Abhinav Upadhyay $<$er.abhinav.upadhyay@gmail.com$>$ \and 
Joerg Sonnenberger $<$joerg@NetBSD.org$>$}
\begin{document}
\maketitle
\section{Abstract}
Unix like operating systems have withstood the test of time for 40 odd years for
several technical and philosophical reasons. One of those reasons is the
quality documentation they ship in the form of man pages. They are a single point
source of consultation for most of our routine work. But there has been a lack of
good search tool to accompany this documentation which would increase their
usefulness manifolds.

Traditionally \textit{apropos} [1] has been there to act as a search
interface but to a large extent it hasn't really lived up to it's expectation.
\textit{apropos} was
developed in the early days of Unix when computing resources were scarce and
which was the primary reason that it's design was kept so
minimal and simple.

In this paper the limitations of the traditional implementation of
\textit{apropos} are discussed along with brief details of how those limitations
were solved in this implementation and also what other modern implementations
are available out there.
\section{Introduction}
The classical version of \textit{apropos} has been implemented by simply
indexing the keywords in the NAME section of the man pages in a plain text file
(whatis.db) [2]
and performing searches on it. The reason for this simple design was most
probably lack of computing resources in the early days. The plain text file
consisting of keywords would hardly take few hundreds of kilo bytes of disk space
and performing search on a plain text file is also not very hard.

This simplified design of \textit{apropos(1)} resulted in limited utility and
usability.The searches are limited to the keywords in the NAME section of the
 keywords. If
only a user knows the exact keywords he gets the right results otherwise most of
the times the searches are full of bogus results or a dead end. In the modern
computing world where hard problems like searching the World Wide Web have been
solved [3] to sufficient degree where most of the data is unstructured, then it
makes perfect sense that we leverage the advancement in technology to solve the
problem of an effective man page search tool, considering the fact that the man
pages consist of structured data.

In present times the machines are powerful and  capable of running the search
algorithms efficiently. Disk
space is also cheap enough that indexing of additional metadata from man pages
can be easily afforded. These are necessary prerequisites for building an
effective search tool.

\section{Limitations of Conventional \textit{apropos(1)}}
In this section results of some sample query searches from
the classical version of \textit{apropos} are shown and their shortcomings are
analyzed.

\subsection{Lack of support for free form queries}
The conventional \textit{apropos} is limited to the keywords used in the NAME
section
of the man pages. If the user specifies keywords relevant to the page which he
is looking but probably those keywords are not used in it's NAME section,
then \textit{apropos(1)} would not return
anything relevant. So in essence the user cannot use free form queries like
those supported by modern search applications.
\lstset{language = sh, caption={No results for free form queries}}
\begin{lstlisting}
$ apropos(1) ``add new user''
add new user: nothing appropriate

$ apropos(1) ``get process status''
get process status: nothing appropriate

$ apropos(1) ``termcap database''
termcap database: nothing appropriate

$ apropos(1) ``signal number to string''
signal number to string: nothing appropriate
\end{lstlisting}

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it has no
natural language support. For example in the following listing it can be seen
that while \textit{apropos} returns the correct result for the query \textit{``make directories''} but it fails when the keyword \textit{``directory''} is
used in place of \textit{``directories''}, although the two words are based on
same root word.
\lstset{language = sh, caption={No support for stems or word with same roots},}
\begin{lstlisting}
$ apropos ``make directories''
mkdir (1) - make directories
$ apropos ``make directory''
make directory: nothing appropriate

$ apropos "upgrading software package"
pkg_add (1) - a utility for installing and upgrading software package
distributions
$ apropos "upgrading software packages"
upgrading software packages: nothing appropriate
\end{lstlisting}

Similarly, it is very common for users to misspell keywords in a query and \textit{apropos} has no support for it either.
\lstset{language = sh, caption={No spelling correction},}
\begin{lstlisting}
$ apropos ``copy strings''
stpcpy, stpncpy, strcpy, strncpy (3) - copy strings
$ apropos ``coppy strings''
coppy strings: nothing appropriate
\end{lstlisting}

\subsection{Unintelligible Output}
Because of the way \textit{apropos} works, sometimes it's output can be
unintelligible and it can be very hard for the user to identify the results which
are relevant to him.
\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},}
\begin{lstlisting}
$ \textit{apropos(1)} power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_class_display_register, pmf_system_suspend, pmf_system_resume,
pmf_system_shutdown, pmf_event_register, pmf_event_deregister,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
\subsection{Other Problems}
Apart from the search related problems there are a few issues related to
the way man pages are handled in NetBSD. The different aliases of the man pages
are stored on the filesystem in the from of hard (or soft) links and these
have to be mentioned in the makefiles explicitly using the MLINKS mechanism. This
approach works fine but it is a mess from maintenance point of view.
It should be possible to fix this by utilizing the index already built and
maintained by \textit{apropos} and related tools, but yet again, the
simplistic implementation of \textit{apropos} does not leave any room for
improvement.

\section{Proposed Solution}
A very simple and straightforward solution is proposed to solve these problems.
The idea is to parse and index
the complete content (or at least most of the content) in the form of an
inverted index [4] and use it to build a full text search
interface. Having an index based on the complete content of the man pages solves
many of the search
related problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows:
\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the keywords defined in the NAME section of the man
pages, the keywords used throughout the body of the page also play a role.
For example queries like \textit{``installing new software package''}
will now get results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This allows us to support more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results.

Similarly along with the inverted index, a dictionary of the keywords frequently
occurring in the corpus of man pages can also be built and used to support
spelling suggestion.

In this implementation, both word stemming and spelling correction in our implementation are used to make the search experience more smooth.
\end{description}
\begin{description}
\item[Bookkeeping of man page metadata] \hfill \\
In this implementation we have gone ahead and also indexed additional metadata
related to the man pages, for example we are indexing an md5 hash of the man
page, the device id, inode number and modification time of the man page file so
as to support incremental updating of the index as new man pages are installed
or the old ones are modified/updated.

Similarly we are also maintaining a separate index of all the aliases of a man
page. This will allow us to get rid of all the hard or symbolic links of the man
pages which are scattered throughout the filesystem, and also help in getting rid
of the MKLINK flags in the makefiles.
\end{description}
\subsection{Tools Used}
The two main tasks that we needed to perform were parsing of the man pages and
building an index of the data, and to do these tasks we have used \textit{libmandoc} [6]
and \textit{sqlite} [7] in our implementation.
\begin{description}
\item[\textit{libmandoc}] \hfill \\
\textit{libmandoc} is a library interface to a validating compiler. It provides
interface to parse and build an AST (Abstract Syntax Tree) of the man page, 
and interface for traversing that tree in order to extract the data from nodes
which are of our interest.
\end{description}
\begin{description}
\item[\textit{sqlite}] \hfill \\
\textit{sqlite} is an embedded relational database management system providing a
relatively small and easy to use C library interface. One of the main reason
for choosing it over the myriad of other possible options was that it provides
in built support for full text search through it's FTS virtual table module [8].
The FTS module can be accessed using pretty much standard SQL syntax, and it is
still flexible enough to allow us to implement our own ranking function as
per our needs. Besides that we have been able to utilize the RDBMS support of
\textit{Sqlite} to store additional metadata in the form of normal database
tables without any hassles.
\end{description}
\section{Implementation Details}
Due to space constraints we cannot go into enough implementation details in this
abstract but we would like to talk a bit about some important components of this
project:
\begin{description}
\item[makemandb(1)] \hfill \\
\textit{makemandb(1)} [9] is the key component of this implementation. It
traverses the filesystem, reads the raw man page source files, feeds them to them
\textit{libmandoc} and then stores the extracted data in the database using
\textit{sqlite}.
\end{description}
\begin{description}
\item[apropos(1)] \hfill \\
This is the version of \textit{apropos(1)} written from scratch utilizing the
full text search functionality of \textit{sqlite}. Unlike the classical version
of \textit{apropos(1)} it only displays the top 10 results relevant to the user
query and most of the times this is sufficient as we shall see later. Also, it
shows a brief snippet for each of the search results, making it more easy to
find out the relevant documents.
\end{description}
\begin{description}
\item[apropos-utils(3)] \hfill \\
\textit{apropos-utils(3)} [10] is a small library interface provided with this
implementation.
It provides functions for querying the FTS index and processing the results in
a user supplied callback function. It's main purpose is to develop different
interfaces on top of this implementation for different use cases. For example we
have built a small CGI application using it for doing the searches from a web
browser, similarly we also built an IRC bot using this 
\end{description}
\section{Results}
We would like to show how our implementation solves the problems cited earlier
with the classical version of \textit{apropos(1)}.
\footnote{Although this implementation of \textit{apropos(1)} returns 10 results
in normal case for a query but in the following listings we have snipped the
results to save space
}
\lstset{language = sh, caption={Add new user},}
\begin{lstlisting}
$ apropos ``add new user''
ssh-add(1)      adds private key identities to the authentication agent
...on the command line. If any file requires a passphrase, ssh-add
asks for the passphrase from the user. The passphrase is read from the
user's tty. ssh-add retries the last passphrase if multiple identity
files are given...

chpass(1)       add or change user database information
add or change user database information

useradd(8)      add a user to the system
The useradd utility adds a user to the system, creating and populating
a home directory if necessary. Any skeleton files will be provided for
the new user if they exist in the skel-dir directory (see the k
option). Default...
.
.
.
\end{lstlisting}
\lstset{language = sh, caption={make directory},}
\begin{lstlisting}
$ apropos ``make directory''
make(1) maintain program dependencies
...CURDIR A path to the directory where make was executed. Refer to
the description of PWD for more details. MAKE The name that make was
executed with argv[0] . For compatibility make also sets .MAKE with
the same value. The...

mkdir(1)        make directories
make directories

ln(1)   make links
...a directory in which to place the link; otherwise it is placed in
the current directory. If only the directory is specified, the link
will be made to the last component of source_file . Given more than
two arguments, ln makes...

mkfifo(1)       make fifos
make fifos...of a=rw mkfifo requires write permission in the parent
directory. mkfifo exits 0 if successful, and >0 if an...

mkdir(2)        make a directory file
...will contain the directory has been exhausted. EDQUOT The user's
quota of inodes on the file system on which the directory is being
created has been exhausted. EIO An I/O error occurred while making the
directory entry or...
\end{lstlisting}
\lstset{language = sh, caption={get process status},}
\begin{lstlisting}
$ apropos ``get process status''
ps(1)   process status
process status

netstat(1)      show network status
...use the fstat(1) command to find out which process or processes
hold references to a socket. The interface display...show network
status

bounce(8)       Postfix delivery status reports
...the master(8) process manager.  The bounce(8) daemon
processes two types of service requests: Append a recipient
(non-)delivery status record to a per message log file. Enqueue a
delivery status notification message, with a...

ocsp(1) Online Certificate Status Protocol utility
Online Certificate Status Protocol utility...it is in client mode. The
request(s) the responder processes can be either specified on the
command line (using...

fstat(1)        display status of open files
...the owner of the process (effective UID). CMD The command name of
the process. PID The process ID. FD The...display status of open files
\end{lstlisting}

\begin{figure}[htp]
\includegraphics[scale=0.30]{/home/abhinav/development/AsiaBSDCon/spell.png}
\caption{The spell corrector and the web interface in action}
\label{}
\end{figure}
\section{Related Work}
There are at least two projects which are related to this in some way.
\begin{description}
\item[man-db] \hfill \\
\textit{man-db} [11] is a complete implementation of the man page documentation
system and it is used on a number of GNU/Linux distributions.\textit{man-db}
takes an interesting approach for indexing the man page data. Unlike the
classical \textit{apropos(1)}, it uses a Berkley DB database but still it's index
is limited to the NAME section only. It adds an option `K' to man(1) to allow a
crude full text search but it is not very efficient nor effective.
\end{description}
\begin{description}
\item[mandocdb] \hfill \\
\textit{mandocdb} [12] is more in line with the goals of our project but it takes
a novel approach. It indexes the keywords extracted from the man pages in a
key-value store using btree(3) [13]. It also comes with it's own implementation
of \textit{apropos(1)} which performs search using this key-value store. The
main key point of this implementation is that it exploits the semantic structure
of the man pages. 
\end{description}
\section{Future Work}
\begin{description}
\item[Work on Ranking Algorithm] \hfill \\
We have used a ranking algorithm based on probabilistic model [14] of information
retrieval. It is essentially based on the Okapi BM25F algorithm [15] and uses
certain parameters whose values are usually dependent on the corpus and the
search application. For example we have used certain weights for different
sections of man pages. At the moment these values have been determined manually
but we would like to use some machine learning techniques in the coming days to
automate this.
\end{description}
\begin{description}
\item[Fix Some Loose Ends In Parsing of man pages] \hfill \\
Although the parsing routine is working fine for most purposes on the man pages
supplied with NetBSD and also on most of the man pages found in pkgsrc. But there
are perhaps a few corner cases still to be fixed. Besides that another issue that
is to be addressed is the parsing of the escape sequences which are used
extensively in the \textit{man(7)} [16] based man pages. We would like to address
these issues in the coming days.
\end{description}
\section{Acknowledgement}
This project was developed as part of \textbf{Google Summer of Code 2011} [17],
so thanks to Google for sponsoring it. We would also like to thank
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. We would also like to thank \textbf{David Young} who was involved with
this project closely and offered useful help and guidance. A special thanks goes
to \textbf{Thomas Klausner} who helped in writing and reviewing our man pages.
Thanks to Petra Zeidler for administering the GSoC program for \textit{The NetBSD
Foundation}.
\begin{thebibliography}{20}
\bibitem{1}
\emph{NetBSD manual page for apropos(1)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?apropos++NetBSD-5.1

\bibitem{2}
\emph{NetBSD manual page for makewhatis(1)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?makewhatis++NetBSD-5.1

\bibitem{3}
Brin, S.; Page L.
\emph{The Anatomy of a Large-Scale Hypertextual Web Search Engine}
Computer Networks and ISDN Systems
30:107-117, 1998
\bibitem{4}
Manning; Raghwan; Schutze
\emph{Introduction to information retrieval},
3-9,
2008
\bibitem{5}
Porter, M. F.
\emph{An algorithm for suffix stripping},
Program,
14(3): 130-137,
1980

\bibitem{6}
\emph{mdocml online manual page for
libmandoc}
\newline http://mdocml.bsd.lv/mandoc.3.html

\bibitem{7}
\emph{Sqlite home page}
\newline http://sqlite.org

\bibitem{8}
\emph{Sqlite FTS3 and FTS4 Extensions}
\newline  http://sqlite.org/fts3.html

\bibitem{9}
\emph{Online manual page for makemandb(1)}
\newline http://netbsd-soc.sourceforge.net/projects/apropos\_replacement/makemandb.html1

\bibitem{10}
\emph{Online manual page for apropos-utils(3)}
\newline http://netbsd-soc.sourceforge.net/projects/apropos\_replacement/apropos-utils.html3

\bibitem{101}
\emph{man-db, the on-line manual database}
\newline http://man-db.nongnu.org/

\bibitem{12}
\emph{Online manual page for mandocdb(1)}
\newline http://mdocml.bsd.lv/mandocdb.8.html

\bibitem{13}
\emph{NetBSD online manual page for btree(3)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?btree+3+NetBSD-5.1

\bibitem{14}
Fuhr, Norbert
\emph{Probabilistic models in information
retrieval}
The Computer Journal
1992
\bibitem{15}
Zaragoza H.; Craswell N.; Taylor M.;
Saria S.; Robertson S.
\emph{Microsoft Cambridge at TREC–13: Web
and HARD tracks}
In proceedings of TREC-2004

\bibitem{16}
\emph{Online manual page for man(7)}
\newline http://mdocml.bsd.lv/man.7.html

\bibitem{17}
\emph{Google Summer of Code home page}
\newline http://code.google.com/soc/

\bibitem{18}
\emph {The mdocml project home page}
\newline http://mdocml.bsd.lv
\end{thebibliography}
\end{document}