### 2012-02-02 apropos_replacement.tex
Fix some typos and rename files.

\documentclass[titlepage, a4paper, 12pt]{article}
\usepackage{listings}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\lstset{ %
language=sh,
basicstyle=\footnotesize,
numbers=none,
backgroundcolor=\color{white},
showspaces=false,
showstringspaces=false,
showtabs=false,
frame=single,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
title=\lstname
}
\title{Apropos Replacement: Development of a full text search tool for man pages \newline
(Initial Outline)}
\author{Abhinav Upadhyay $<$er.abhinav.upadhyay@gmail.com$>$ \and 
Joerg Sonnenberger $<$joerg@NetBSD.org$>$}
\begin{document}
\maketitle
\section{Abstract}
Unix like operating systems have withstood the test of time for 40 odd years for
several technical and philosophical reasons. One of those reasons is the
quality documentation they ship in the form of man pages, which are a single
source of consultation for most of our routine work. However, so far, there has been a lack of
good search tool to accompany this documentation, which would increase their
usefulness manifolds.

Traditionally, \textit{apropos} [1] has been there to act as a search
interface but to a large extent it hasn't really lived up to its expectation.
\textit{apropos} was
developed in the early days of Unix when computing resources were scarce and
that is the primary reason for its minimalistic and simple design. The biggest
limitation of \textit{apropos} is its limited search capabilities.

This paper discusses a new implementation of \textit{apropos} which was done
as part of \emph{Google Summer of Code 2011} [17]. The goal of this project was
to replace the conventional apropos in NetBSD with a modern version which
supports full text searches.
\section{Introduction}
The classical version of \textit{apropos} has been implemented by simply
indexing the keywords in the NAME section of the man pages in a plain text file
(whatis.db) [2]
and performing searches on it. The reason for this simple design was most
probably lack of computing resources in the early days. The plain text file
consisting of keywords hardly takes few hundreds of kilo bytes of disk space
and performing search on a plain text file is quite easy.

This simplified design of \textit{apropos} resulted in limited utility and
usability, as the searches are limited to the keywords defined in the NAME
section of the man pages. Only if a user knows the exact keywords do they get
the appropriate results; otherwise most of
the times the searches are full of irrelevant results or possibly a dead end. 
In the modern
computing world where hard problems like searching the World Wide Web have been
solved [3] to a sufficient degree, it makes perfect sense to leverage the
advancement in technology in order to solve the problem. This problem is made
even simpler considering that the man pages consist of structured data.

In present times, the machines are powerful and  capable of running the search
algorithms efficiently. Disk
space is also sufficiently cheap that any extra space incurred by indexing of
additional metadata from man pages
can be easily afforded. These are necessary prerequisites for building an
effective search tool.

In this paper, the limitations of the traditional implementation of apropos
are discussed along with brief details of how these were overcome as part of
this project. Additionally, we compare and contrast our approach with other
modern implementations available out there.

\section{Limitations of Conventional \textit{apropos}}
In this section results of some sample query searches from
the classical version of \textit{apropos} are shown and their shortcomings are
analyzed.

\subsection{Lack of support for free form queries}
As noted earlier the conventional \textit{apropos} is limited to the keywords
used in the NAME section of the man pages. This means that the users do not have
a whole lot of choice for specifying keywords in their queries. Besides they
cannot search for keywords which are not usually specified in the title of a man
page, for example a query like ``EINVAL'' would return no results.
\lstset{language = sh, caption={No results for free form queries}}
\begin{lstlisting}
$ apropos ``add new user''
add new user: nothing appropriate

$ apropos ``get process status''
get process status: nothing appropriate

$ apropos ``termcap database''
termcap database: nothing appropriate

$ apropos ``signal number to string''
signal number to string: nothing appropriate
\end{lstlisting}

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it is not smart enough to provide basic natural language processing constructs,
like stemming or spelling corrections.
For example in the following listing it can be seen
that while \textit{apropos} returns the correct result for the query
\textit{``make directories''}, it fails when the keyword \textit{``directory''}
is used in place of \textit{``directories''}, even though the two words are based
on same root word.
\lstset{language = sh, caption={No support for stems or words with same roots},}
\begin{lstlisting}
$ apropos ``make directories''
mkdir (1) - make directories
$ apropos ``make directory''
make directory: nothing appropriate

$ apropos ``upgrading software package''
pkg_add (1) - a utility for installing and upgrading software package
distributions
$ apropos ``upgrading software packages''
upgrading software packages: nothing appropriate
\end{lstlisting}

Similarly, it is very common for users to misspell keywords in a query and \textit{apropos} has no support for it either.
\lstset{language = sh, caption={No spelling correction},}
\begin{lstlisting}
$ apropos ``copy strings''
stpcpy, stpncpy, strcpy, strncpy (3) - copy strings
$ apropos ``coppy strings''
coppy strings: nothing appropriate
\end{lstlisting}

\subsection{Unintelligible Output}
Because of the way \textit{apropos} works, sometimes its output can be
unintelligible and it can be very hard for the user to identify relevant results.

\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},}
\begin{lstlisting}
$ apropos power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
\subsection{Other Problems}
Apart from the search related problems there are a few issues related to
the way man pages are handled in NetBSD. The different aliases of the man pages
are stored on the file system in the from of hard (or soft) links and these
have to be mentioned in the makefiles explicitly using the MLINKS mechanism. This
approach works fine but it is a mess from maintenance point of view.
It should be possible to fix this by utilizing the index already built and
maintained by \textit{apropos}, but yet again, the simplistic implementation of
\textit{apropos} does not leave any room for improvement.

\section{Proposed Solution}
A very simple and straightforward solution is proposed to solve these problems as
part of this project.
The idea is to parse and index
the complete content (or at least most of the content) in the form of an
inverted index [4] and use it to build a full text search
interface. Having an index based on the complete content of the man pages solves
many of the search
related problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows:
\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the the NAME section.
For example, queries like \textit{``installing new software package''} will
now produce relevant results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation, the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This enables support for more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results.

Similarly along with the inverted index, a dictionary of the keywords frequently
occurring in the corpus of man pages is also built and used to support
spelling suggestion.
\end{description}
\begin{description}
\item[Bookkeeping of man page metadata] \hfill \\
In this implementation, additional metadata related to the man pages, for example
an md5 hash, the device id, inode number and modification time
of the man page files are indexed and stored so
as to support fast and hassle free update of the index as new man pages are
installed or the old ones are modified/updated.

Similarly a separate index of all the man page aliases is stored and maintained.
This provides an option to get rid of all the hard or symbolic links of the man
pages scattered throughout the filesystem, and also for clean up of the MLINKS
mess in the makefiles.
\end{description}
\subsection{Tools Used}
The two main operations that are critical for this project are parsing of the
man pages and building of an index of the
data obtained from the parsing process. \textit{libmandoc} [6]
and \textit{sqlite} [7] have been used for performing these the two tasks.
\begin{description}
\item[\textit{libmandoc}] \hfill \\
\textit{libmandoc} is a library interface to a validating compiler. It provides
interface to parse and build an AST (Abstract Syntax Tree) of the man page. It
also provides an interface for traversing that tree in order to extract the data
from its nodes.
\end{description}
\begin{description}
\item[\textit{sqlite}] \hfill \\
\textit{sqlite} is an embedded relational database management system providing a
relatively small and easy to use C library interface. One of the main reasons
for choosing it over the myriad of other possible options is that it provides
built in support for full text search through its FTS virtual table module [8].
The FTS module can be accessed using pretty much standard SQL syntax, and it is
still flexible enough to accept user supplied ranking function to suit the needs
of the application. Besides that, another advantage of \textit{sqlite} is that, 
being an RDBMS, it makes it very easy to store additional metadata in the form
of normal database tables without any hassles.
\end{description}
\section{Implementation Details}
Due to space constraints it is not possible to go into enough implementation
details, the most important components of this project are described in brief
in this section.
\begin{description}
\item[makemandb] \hfill \\
\textit{makemandb} [9] is the key component of this implementation. It is a
command line tool which
traverses the filesystem, reads the raw man page source files, feeds them to the
\textit{libmandoc} parser and then stores the extracted data in the database
using \textit{sqlite}.
\end{description}
\begin{description}
\item[apropos] \hfill \\
This is the version of \textit{apropos} written from scratch utilizing the
full text search functionality of \textit{sqlite}. Unlike the classical version
of \textit{apropos}, it only displays the top 10 results relevant to the user
query and most of the times this is sufficient as will be seen later. Also, it
shows a brief snippet of matching text for each of the search results, making it
more easy to identify the relevant documents.
\end{description}
\begin{description}
\item[apropos-utils] \hfill \\
\textit{apropos-utils} [10] is a small library interface provided with this
implementation. It provides functions for querying the FTS index and for
processing the results in a user supplied callback function. Its main purpose
is to develop different interfaces on top of it for different
use cases. For example a small CGI application was built using it for doing
the searches from a web browser, similarly an IRC bot was also developed utilizing this interface.
\end{description}
\section{Results}
This section shows results of some of the sample queries on the version of
developed as part of this project
\textit{apropos} to demonstrate how it solves many of the problems cited earlier
with the classical version of \textit{apropos}.
\footnote{Although this implementation of \textit{apropos} returns 10 results
in normal cases for a query but in the following listings the output has been
snipped to save space
}
\lstset{language = sh, caption={Add new user},}
\begin{lstlisting}
$ apropos ``add new user''
ssh-add(1)      adds private key identities to the authentication agent...on the command line. If any file requires a passphrase, ssh-add asks for the passphrase from the user. The passphrase is read from the user's tty. ssh-add retries the last passphrase if multiple identity files are given...

chpass(1)       add or change user database information
add or change user database information

useradd(8)      add a user to the system
The useradd utility adds a user to the system, creating and populating a home directory if necessary. Any skeleton files will be provided for the new user if they exist in the skel-dir directory (see the k option). Default...
.
.
.
\end{lstlisting}
\lstset{language = sh, caption={make directory},}
\begin{lstlisting}
$ apropos ``make directory''
make(1) maintain program dependencies
...CURDIR A path to the directory where make was executed. Refer to the description of PWD for more details. MAKE The name that make was executed with argv[0] . For compatibility make also sets .MAKE with the same value. The...

mkdir(1)        make directories
make directories

ln(1)   make links
...a directory in which to place the link; otherwise it is placed in the current directory. If only the directory is specified, the link will be made to the last component of source_file . Given more than two arguments, ln makes...

mkfifo(1)       make fifos
make fifos...of a=rw mkfifo requires write permission in the parent directory. mkfifo exits 0 if successful, and >0 if an...

mkdir(2)        make a directory file
...will contain the directory has been exhausted. EDQUOT The user's quota of inodes on the file system on which the directory is being created has been exhausted. EIO An I/O error occurred while making the directory entry or...
\end{lstlisting}
\lstset{language = sh, caption={signal number to string},}
\begin{lstlisting}
$ apropos ``signal number to string''
psignal(3)      system signal messages
...the signal number is not recognized sigaction(2) , the string Unknown signal is produced. The psiginfo function produces the same output as the psignal function, only it uses the signal number information from the si argument. The message strings can...

intro(2)        introduction to system calls and error numbers
...undefined signal to a signal(3) or kill(2) function). 23 ENFILE Too many open files in system . Maximum number...shell. Pathname A path name is a NUL -terminated character string starting with an optional slash \&/ , followed by zero or...

dump_lfs(8)     filesystem backup
...low estimates of the number of blocks to write, the number of tapes it...a string containing embedded formatting commands for strftime(3). The total formatted string is...in. If dump_lfs rdump_lfs receives a SIGINFO signal (see the status argument of stty...
\end{lstlisting}

In \textbf{Figure 1} the spell corrector can be seen in action along with the web
interface that was developed using \textit{apropos-utils}.
\begin{figure}[htp]
\includegraphics[scale=0.30]{/home/abhinav/development/AsiaBSDCon/spell.png}
\caption{The spell corrector and the web interface in action}
\label{}
\end{figure}

\section{Related Work}
There are at least two projects which are related to this in some way.
\begin{description}
\item[man-db] \hfill \\
\textit{man-db} [11] is a complete implementation of the man page documentation
system and it is used on a number of GNU/Linux distributions. \textit{man-db}
takes an interesting approach for indexing the man page data. Unlike the
classical \textit{apropos}, it uses a Berkley DB database but still its index
is limited to the NAME section only. It adds an option `K' to \textit{man(1)} to
allow a crude full text search but it is not very efficient nor effective.
\end{description}
\begin{description}
\item[mandocdb] \hfill \\
\textit{mandocdb} [12] is more in line with the goals of this project but it
takes a novel approach. It indexes the keywords extracted from the man pages in a
key-value store using btree(3) [13]. It also comes with its own implementation
of \textit{apropos} which performs search using this key-value store. The
main key point of this implementation is that it exploits the semantic structure
of the man pages.
\end{description}
\section{Future Work}
\begin{description}
\item[Work on Ranking Algorithm] \hfill \\
A ranking algorithm based on probabilistic model of information retrieval [14]
has been implemented to perform relevance ranking of the search
results. It is essentially based on the Okapi BM25F algorithm [15] and
uses certain parameters whose values are usually dependent on the corpus and the
search application. For example, certain weight parameters have been assigned to
different sections of man pages, signifying how important the information in a
particular section is as compared to the rest of the sections. At the moment
these values have been determined
manually but one goal is to use some supervised machine learning techniques in
order to automate this.
\end{description}
\begin{description}
\item[Fix Some Loose Ends In Parsing of man pages] \hfill \\
Although the parsing routine is working fine for most purposes on the man pages
supplied with NetBSD and also on most of the man pages found in Pkgsrc. But there
are perhaps a few corner cases yet to be identified and fixed. Besides that,
another issue that is to be addressed is the parsing of the escape sequences
which are used extensively in the \textit{man(7)} [16] based man pages. These
issues are to be addressed in the coming days.
\end{description}
\section{Availability}
The code for this project is hosted on \textit{Github} and it is under active development. It can be obtained from:\newline
 \emph{https://github.com/abhinav-upadhyay/apropos\_replacement}
\section{Acknowledgement}
This project has been developed as part of \emph{Google Summer of Code 2011}
[17], so thanks to \textbf{Google} for sponsoring it. Special thanks to
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. We would like to thank \textbf{David Young} who was involved with
this project closely and offered useful help and guidance throughout. A special
thanks goes to \textbf{Thomas Klausner} who helped in writing and reviewing
man pages for this project. Thanks to \textbf{Petra Zeidler} for administering
the GSoC program for \textit{The NetBSD Foundation}.
\begin{thebibliography}{20}
\bibitem{1}
\emph{NetBSD manual page for apropos(1)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?apropos++NetBSD-5.1

\bibitem{2}
\emph{NetBSD manual page for makewhatis(1)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?makewhatis++NetBSD-5.1

\bibitem{3}
Brin, S.; Page L.
\emph{The Anatomy of a Large-Scale Hypertextual Web Search Engine}
Computer Networks and ISDN Systems
30:107-117, 1998
\bibitem{4}
Manning; Raghwan; Schutze
\emph{Introduction to information retrieval},
3-9,
2008
\bibitem{5}
Porter, M. F.
\emph{An algorithm for suffix stripping},
Program,
14(3): 130-137,
1980

\bibitem{6}
\emph{mdocml online manual page for
libmandoc}
\newline http://mdocml.bsd.lv/mandoc.3.html

\bibitem{7}
\emph{Sqlite home page}
\newline http://sqlite.org

\bibitem{8}
\emph{Sqlite FTS3 and FTS4 Extensions}
\newline  http://sqlite.org/fts3.html

\bibitem{9}
\emph{Online manual page for makemandb(1)}
\newline http://netbsd-soc.sourceforge.net/projects/apropos\_replacement/makemandb.html1

\bibitem{10}
\emph{Online manual page for apropos-utils(3)}
\newline http://netbsd-soc.sourceforge.net/projects/apropos\_replacement/apropos-utils.html3

\bibitem{101}
\emph{man-db, the on-line manual database}
\newline http://man-db.nongnu.org/

\bibitem{12}
\emph{Online manual page for mandocdb(1)}
\newline http://mdocml.bsd.lv/mandocdb.8.html

\bibitem{13}
\emph{NetBSD online manual page for btree(3)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?btree+3+NetBSD-5.1

\bibitem{14}
Fuhr, Norbert
\emph{Probabilistic models in information
retrieval}
The Computer Journal
1992
\bibitem{15}
Zaragoza, H.; Craswell, N.; Taylor, M.;
Saria, S.; Robertson, S.
\emph{Microsoft Cambridge at TREC–13: Web
and HARD tracks}
In proceedings of TREC-2004

\bibitem{16}
\emph{Online manual page for man(7)}
\newline http://mdocml.bsd.lv/man.7.html

\bibitem{17}
\emph{Google Summer of Code home page}
\newline http://code.google.com/soc/

\bibitem{18}
\emph {The mdocml project home page}
\newline http://mdocml.bsd.lv
\end{thebibliography}
\end{document}