### 2012-02-28 apropos_replacement.tex
Make style consistent with Usenix. Add more details for submission to AsiaBSDCon.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}
\usepackage{listings}
\usepackage{color}
\usepackage{times}
\usepackage{algpseudocode}
%\usepackage[pdftex]{graphicx}
\lstset{ %
language=sh,
basicstyle=\footnotesize,
numbers=none,
backgroundcolor=\color{white},
showspaces=false,
showstringspaces=false,
showtabs=false,
frame=single,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
title=\lstname
}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Apropos Replacement : Development of a full text search tool for man pages}

%for single author (just remove % characters)
\author{
{\rm Abhinav Upadhyay}\\
$<$er.abhinav.upadhyay@gmail.com$>$
\and
{\rm Joerg Sonnenberger}\\
$<$joerg@NetBSD.org$>$
} % end author

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
Manual pages are a key component of Unix like operating systems, they have
played a significant role in the success of Unix over the years. Man pages are
the most authentic source of reference for of a system administrator or a
programmer. However, so far, there has been a lack of good search tool to make
man pages more accessible. \\

Traditionally, \textit{apropos(1)} has been there as a search interface but it
was developed in the early days of Unix when computing resources were scarce
and that is the primary reason for its simple design and limited search
capabilities. \\

This paper discusses a new implementation of \textit{apropos} done by
Abhinav Upadhyay as part of
Google Summer of Code 2011. The goal of this project was to replace the
conventional apropos in NetBSD with a modern version which supports full text
searches. \\

\section{Introduction}
The classical version of \textit{apropos} has been implemented by simply
indexing the keywords in the NAME section of the man pages in a plain text file
(whatis.db) [2]
and performing searches on it. The reason for this simple design was most
probably lack of computing resources in the early days. The plain text file
consisting of keywords hardly takes few hundreds of kilo bytes of disk space
and performing search on a plain text file is quite easy. \\

This simplified design of \textit{apropos} resulted in limited utility and
usability, as the searches are limited to the keywords defined in the NAME
section of the man pages. Only if the users know the exact keywords do they get
the appropriate results; otherwise most of
the times the searches are full of irrelevant results or possibly a dead end. 
In the modern
computing world where hard problems like searching the World Wide Web have been
solved [3] to a sufficient degree, it makes perfect sense to leverage the
advancement in technology in order to solve the problem. This problem is made
even simpler considering that the man pages consist of structured data. \\

In present times, the machines are powerful and  capable of running the search
algorithms efficiently. Disk
space is also sufficiently cheap that any extra space incurred by indexing of
additional metadata from man pages
can be easily afforded. These are necessary prerequisites for building an
effective search tool. \\

In this paper, the limitations of the traditional implementation of apropos
are discussed along with brief details of how these were overcome as part of
this project. Additionally, a comparison of other modern implementations of apropos has been done. \\

\section{Limitations of Conventional \textit{apropos}}
In this section results of some sample query searches from
the classical version of \textit{apropos} are shown and their shortcomings are
analyzed. \\

\subsection{Lack of support for free form queries}
As noted earlier the conventional \textit{apropos} is limited to the keywords
used in the NAME section of the man pages. This means that the users do not have
a whole lot of choice for specifying keywords in their queries. Besides they
cannot search for keywords which are not usually specified in the title of a man
page, for example a query like ``EINVAL'' would return no results. \\

{\tt \small
\lstset{language = sh, caption={No results for free form queries}}
\begin{lstlisting}
$ apropos ``add new user''
add new user: nothing appropriate

$ apropos ``get process status''
get process status: nothing appropriate

$ apropos ``termcap database''
termcap database: nothing appropriate
\end{lstlisting}
}

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it is not smart enough to provide basic natural language processing constructs,
like stemming or spelling corrections.
For example in the following listing it can be seen
that while \textit{apropos} returns the correct result for the query
\textit{``make directories''}, it fails when the keyword \textit{``directory''}
is used in place of \textit{``directories''}, even though the two words are based
on same root word. \\

{\tt \small
\lstset{language = sh, caption={No support for stems or words with same root},}
\begin{lstlisting}
$ apropos ``make directories''
mkdir (1) - make directories
$ apropos ``make directory''
make directory: nothing appropriate

$ apropos ``upgrading package''
pkg_add (1) - a utility for installing and upgrading software package
distributions

$ apropos ``upgrading packages''
upgrading packages: nothing appropriate
\end{lstlisting}
}

Similarly, it is very common for users to misspell keywords in a query and \textit{apropos} has no support for it either.

{\tt \small
\lstset{language = sh, caption={No spelling correction},}
\begin{lstlisting}
$ apropos ``copy strings''
stpcpy, stpncpy, strcpy, strncpy (3) - copy
strings
$ apropos ``coppy strings''
coppy strings: nothing appropriate
\end{lstlisting}
}

\subsection{Unintelligible Output}
Because of the way \textit{apropos} works, sometimes its output can be
unintelligible and it can be very hard for the user to identify relevant results. \\

{\tt \small
\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},}
\begin{lstlisting}
$ apropos power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
}

\subsection{Other Problems}
Apart from the search related problems there are a few issues related to
the way man pages are handled in NetBSD. The different aliases of the man pages
are stored on the file system in the from of hard (or symbolic) links and these
have to be specified in the makefiles explicitly using the MLINKS mechanism. This
approach works fine but it is a mess from maintenance point of view.
It should be possible to fix this by utilizing the index already built and
maintained by \textit{apropos}, but yet again, the simplistic implementation of
\textit{apropos} does not leave any room for improvement.

\section{Solution: A new apropos}
This project proposes a very simple and straightforward solution to solve these
problems and make man pages more usable. The idea is to exploit the inherent
semantics and structure of the man pages. The \textit{mdoc(7)}
macros are semantically quite rich, those macros combined with the context of the surrounding section or context in which they have been used, can be used
as an indicator of the relevancy of the content. For example, the
{\tt \small
.Nm
}
macro when used inside of the
{\tt \small
NAME
}
section holds more significance than if used inside of any other section. Similarly the description of various \textit{errno} values in the
\textit{intro(2)} man page has more important information than the mention of
those \textit{errno} values in the
{\tt \small
RETURN VALUES
}
section of various man pages from section 3. In essence, it should be possible
to parse and extract the structurally meaningful data
out of man pages and index them in the form of an inverted index [4]. An
inverted index is a specialized data structure for building search engines. The index combined with a ranking algorithm which exploits the structured nature of the data to give back more relevant results to the user, can be used to build a
high quality search interface for manual pages.
Such a search tool will solve many of the problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows: \\

\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the the NAME section.
For example, queries like \textit{``installing new software package''} may
now produce relevant results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation, the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This enables support for more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results. \\

Similarly along with the inverted index, a dictionary of the keywords frequently
occurring in the corpus of man pages is also built and used to support
spelling suggestion.
\end{description}
\begin{description}
\item[Bookkeeping of man page metadata] \hfill \\
In this implementation, additional metadata related to the man pages, for example
an md5 hash, the device id, inode number and modification time
of the man page files are indexed and stored.
This allows for very fast and hassle free update of the index as new man pages are installed or the old ones are updated. \\

Similarly a separate index of all the man page aliases is stored and maintained.
This provides an option to get rid of all the hard or symbolic links of the man
pages scattered throughout the filesystem, and also for clean up of the MLINKS
mess in the makefiles.
\end{description}

\subsection{Tools Used}
The two main operations that are critical for this project are parsing of the
man pages and building of an index of the
data obtained from the parsing process. These are highly specific problems, and
solving them from scratch would be a separate project in its own right, therefore this project tried to avoid reinventing the wheel and used existing
tools and libraries which are battle tested and very successful. \textit{libmandoc} [6] from \textit{mdocml} [18] project has been used for parsing the man pages and \textit{sqlite} [7] for indexing and storing the parsed
data on the file system.

\begin{description}
\item[\textit{libmandoc}] \hfill \\
\textit{libmandoc} is a library interface to a validating compiler for man pages. It provides
interface to parse and build an AST (Abstract Syntax Tree) of the man page. It
also provides an interface for traversing that tree in order to extract the data
from its nodes.
\end{description}

\begin{description}
\item[\textit{sqlite}] \hfill \\
\textit{sqlite} is an embedded relational database management system providing a
relatively small and easy to use C library interface. One of the main reasons
for choosing it over the myriad of other possible options is that it provides
built in support for full text search through its FTS virtual table module [8].
The FTS module can be accessed using pretty much standard SQL syntax, and it is
still flexible enough to accept user supplied ranking function to suit the needs
of the application. Besides that, another advantage of \textit{sqlite} is that, 
being an RDBMS, it makes it very easy to store additional metadata in the form
of normal database tables without any hassles.
\end{description}

\section{Implementation Details}
Due to space constraints it is not possible to go into enough implementation
details, the most important components of this project are described in brief
in this section.

\subsection{makemandb}
\textit{makemandb(8)} [9] is the key component of this implementation. It is a
command line tool which traverses the filesystem, reads the raw man page source files, extracts the interesting data from them using the \textit{libmandoc} parser and then stores that data in an FTS table using \textit{sqlite}. \\

\begin{description}
\item[Operation of makemandb] \hfill \\
\textit{makemandb} uses a two pass algorithm to index the man pages. In the first pass it obtains the list of directories containing man page files from
\textit{man.conf} and starts traversing those directories. It obtains the
\textit{\{dev\_t, ino\_t, mtime\}} of all the files using \textit{stat(2)}.
A similar set of metadata is stored in the database (referring to the last successful indexing operation), \textit{makemandb} takes a difference (the set difference operation) of the two sets of metadata to obtain the list of man page files on the file system which are either newly installed (new \textit{\{dev\_t, ino\_t\}} pairs) or they have been updated (changed \textit{mtime}). \\

In the second pass \textit{makemandb} generates md5 checksum of the files obtained from the previous stage and for each of these checksums, it checks in
the db, whether this md5 checksum already exists in the db or not. If the md5 checksum is already present in the database, then this was just a false alarm and \textit{makemandb} does not need to parse the file, it simply updates the metadata of that file in the database. In case the md5 checksum of a file does not already exist in the database, then that means the file is either a newly installed man page or an older copy of a man page was updated,
\textit{makemandb} feeds such man pages to the parser and extracts interesting
data out of them, later on storing the data in the FTS table in the database.
\end{description}

\begin{description}
\item[Database Schema] \hfill \\
During the initial stage of the project, the structure of the database was kept
simple. The FTS table consisted of 3 columns:
{\tt
name, name\_desc, desc
}
The
{\tt
name
}
and
{\tt
name\_desc
}
columns stored the name of the man page and the one line description from the
{\tt
NAME
}
section, while the
{\tt
desc
}
column stored the content from the rest of the sections. This approach worked
quite well for a small set of man pages (for example the 4000 or so man pages in
the base set of NetBSD). However, as the number of documents were increased (for example including man pages from pkgsrc), the quality of search started deteriorating. This was mainly because bulk of the parsed data was being stored
in the same column in the database and the ranking algorithm had no way of
identifying more relevant results apart from the number of matches in any page. \\

To rectify this, first there was an attempt to try out better ranking schemes
(the various ranking schemes that were tried out during the development of the
project are discussed later). Better ranking schemes required storing precomputed
term weights [19] on the file system, this almost doubled the storage requirements, therefore such ranking schemes were discarded later on, and the
database schema was decomposed further into more columns. The decomposition into
more columns was done to represent the most commonly occurring sections in usual Unix man pages, for example
{\tt NAME, DESCRIPTION, LIBRARY, RETURN VALUES, EXIT STATUS, ERRORS}
etc. Such a decomposition allowed for an elaborate ranking algorithm, which could
assign different weights to each of these columns, based on their relevancy. The
final schema of the database is shown in {\bf Listing 5}
\end{description}

\subsection{apropos}
\textit{apropos} was written from scratch to use the FTS index created by
\textit{makemandb}. Since, unlike the traditional \textit{apropos}, this version
of \textit{apropos} is not limited to searching within only the {\tt NAME}
section, the resulting
number of search results is usually quite high, therefore new \textit{apropos}
employs a
sophisticated ranking algorithm to filter out the most relevant results and rank
them up. To avoid cluttering the output and make the user interface clean, the
new \textit{apropos} displays only the top 10 results by default with options
to display more results if required. During the development and testing of the
project it has been observed that, for a reasonable query, the top 10 results
are most often sufficient. \\

Another key aspect of any search application is to filter out or avoid stop-words
. Words like \textit{``the'', ``a'', ``an'', ``'this''} etc. come under the
category of stop-words, words which are \emph{very} commonly used in the
language. Stop-words usually skew the search results because of their sheer
frequency in the corpus. The usual practice in the information retrieval world
is to filter such stop-words while building the index, however, in case of this
project, it proved to be non-trivial within the given time constraints.
Therefore it was decided that, rather than eliminating stop-words from the index,
a reasonable trade off can be made by filtering out any stop-words from the
user's search query and querying the database for only the keywords which do
not come under the category of stop-words. \\

{\tt \small
\lstset{language = XML, caption={Database Schema},}
\begin{lstlisting}
The Database has three tables at present:

(1) mandb:
    The main FTS table which contains all
    the content parsed from man pages

  COLUMN NAME       DESCRIPTION
  1. section        Section number
  2. name           The name of the page  
  3. name_desc      Short description from
                    NAME section  
  4. desc           Contains the content
                    from DESCRIPTION
                    section and any other
                    section which is not
                    stored in a separate
                    column.
  5. lib            The LIBRARY section
  6. return_vals    RETURN VALUES section
  7. env            ENVIRONMENT section
  8. files          FILES section
  9. exit_status    EXIT STATUS section
  10. diagnostics   DIAGNOSTICS section
  11. errors        ERRORS section
  12. md5_hash      An md5 hash of the
                    contents of the man
                    page (UNIQUE).
(2) mandb_meta:
    This table maintains essential
    metadata of all the indexed
    man page files.

  COLUMN NAME       DESCRIPTION
  1. device         (dev_t)Logical device
                    number from stat(2)
  2. inode          (ino_t)Inode number
                    from stat(2)
  3. mtime          Last modification time
                    from stat(2)
  4. file           Absolute path name
                    (UNIQUE constraint)
  5. md5_hash       MD5 Hash of the man
                    page
  6. id             A unique integer ID
                    referring to a page
                    in mandb (PRIMARY KEY)

(3) mandb_links:
    This is an index of all the hard/soft
    links of the man pages. The list of
    the linked pages is usually obtained 
    from the multiple .Nm entries in the
    page.

  COLUMN NAME       DESCRIPTION
  1. link           The name of the hard/
                    soft link (UNIQUE index) 
  2. target         Name of the target
                    page to which it
                    points  
  3. section        The section number  
  4. machine        The machine
                    architecture (if any)
                    for which the page is
                    relevant.
\end{lstlisting}
}

\subsection{apropos-utils}
\textit{apropos-utils} [10] is a small library interface provided with this
implementation. It provides functions for querying the FTS index and for
processing the results in a user supplied callback function. Its main purpose
is to develop different interfaces on top of it for different
use cases. For instance a CGI front end has been built using it for doing
the searches from a web browser, similarly an IRC bot was also developed utilizing this interface.

\subsection{Ranking Algorithm}
The ranking algorithm is the most interesting and most crucial component of
any search application, it is the deciding factor of the usefulness of the
application. This project took several stabs at coming up with a suitable ranking
algorithm for \textit{apropos}.

\begin{description}
\item[1.] \hfill \\
Initially the database schema was quite simple, it consisted of only three
columns
{\tt 
name, name\_desc, desc
}
(as described previously). At this point, there weren't too many documents in
the corpus, therefore nothing sophisticated was required. The basic intuition was
that, a match found in the
{\tt NAME} section of a man page has more weight as compared to a match found
in any other section. This ranking scheme worked well for a small set of man
pages in the corpus.
\end{description}

\begin{description}
\item[2.] \hfill \\
As more and more man pages were being added to the corpus for testing purposes,
the quality of search results started to dilute. At this point of time, a
completely new approach was adopted for ranking the results. In the literature
related to information retrieval, \textit{tf-idf} [19] weights based ranking models are very common. The \textit{tf} in \textit{tf-idf} stands for \emph{Term Frequency} and \textit{idf} stands for \emph{Inverse Document Frequency}. \\

{\bf Term Frequency}: Term frequency refers to the count of the number of times
a given term
{\tt \small
t
}
appears in a given document
{\tt \small
d
} \\

{\bf Inverse Document Frequency:} The \textit{idf} of a term
{\tt \small
t
}
is the number of documents in which the term appears \textit{at least once}. \\

Term frequency is a \textit{local} factor, it is concerned only with the number
of occurrences of the search terms in one particular document at a time.
While inverse document frequency is a \textit{global} factor, in the sense that,
it indicates the discriminating power of a term. If a term appears in only a
selected set of documents, then that term separates those set of
documents from the rest. Therefore, ranking obtained by combining these two factors
brings up more relevant documents.

The weight of a term
{\tt \small t} in document {\tt \small d} is calculated by the following
formula:

\begin{center}
\framebox{\parbox[b]{2 in}{$weight  = tf \times idf$}}
\end{center}
Where \textit{tf} = Term frequency of term {\tt \small t} in document
{\tt \small d} \\

Inverse document frequency is calculated using the following formula:
\begin{center}
\framebox{\parbox[b]{2 in}{$idf = log (\frac{N}{N_t})$}}
\end{center}
Where $N$ = Total number of documents in the corpus \\
$N_t$ = Number of documents in which term $t$ occurs (at least once). \\

So for a term which appears in only one document it will have
$idf = log(n)$,
while a term which appears in all the documents, it will have
$idf = log(1) = 0$

For example a term like \textit{``the''} will have a high
\textit{term frequency} in any
document, but at the same time it will have a lower \textit{inverse document frequency} (almost close to $0$), which will nullify its effect on the quality of search results.

A lot of research has been done on \textit{tf-idf} weights based ranking
algorithms and various techniques have been proposed over the years. According
to a survey done by \textit{Salton} and \textit{Buckley} in 1988 [21], the
following ranking scheme was found to be most effective:

\begin{center}
\framebox{\parbox[b]{2 in}{$\frac{tf \cdot log(\frac{N}{N_t})}{\sqrt{\sum{(tf \cdot log(\frac{N}{N_t}))^2}}}$}}
\end{center}

The implementation of the above scheme proved to be \emph{very} slow. It required
too many computations and all of these were required to be done for each
search result separately to compute its relevance weight. In an state of the
art search application, the index would usually store precomputed \textit{tf-idf}
weights on the file system, in which case the overhead of computing weights of
each individual search result would require very less work, leading to better
performance. An attempt was made to compute the \textit{tf-idf} weights while
indexing the man pages and store them in a separate table. This lead to high
quality search results but it also brought up a serious issue. Storing
precomputed weights on the file system, more than doubled the storage
requirements of the database index. As a concrete example, it took close
to $90 MB$ to index about $8000$ documents. Without the precomputed weights, 
the database size requirement was close to $45 MB$ for the same number of pages.
\end{description}

\begin{description}
\item[3.] \hfill \\
Since the conventional \textit{apropos} required only a few hundred kilobytes of
disk space, the disk space requirements of new \textit{apropos} seemed
hyperbolic, therefore it was decided later on to look for alternative ranking
schemes. At this point the database schema was decomposed into several columns
representing some of the most common sections found in man pages
(see {\bf Listing 5}).

At this point a fast yet sophisticated ranking algorithm was adopted.
\emph{OKAPI BM25F} [15] is considered one of the most successful probabilistic
models for information retrieval. \emph{Okapi BM25F} is based on \emph{tf-idf}
scheme but it takes into account for the structure of the data. It allows for
separate weights to be assigned for different parts of a structured document to
come up with better relevance rankings. This scheme works very well for man pages
because of their highly structured content. The algorithm is described in pseudo-code below. \textit{apropos} calls this function for each of the document
obtained as part of the result set for the query and computes its relevancy
weight. The documents are ranked in decreasing order or the weights returned
from this function. \\

In the following algorithm: \\
{\tt
\lstset{language = sh, caption={},}
\begin{lstlisting}
nhitcount =  Number of occurrences of
             phrase p in column c in
             current document

nglobalhitcount = Number of occurrences of
                  phrase p in column c in
                  all the documents.

ndocshitcount = Number of documents in
                which phrase p occurs
                in column c, at least
                once.
\end{lstlisting}
}

\begin{algorithmic}
\State ${\tt tf} \gets 0.0$
\State ${\tt idf} \gets 0.0$
\State ${\tt k} \gets 3.5$
\Comment 
{k is an experimentally determined parameter}
\State ${\tt doclen} \gets $ length of the current document
\State ${\tt ndoc} \gets $ Total number of documents in the corpus
\For {each phrase {\tt p} in the user query}
\For {each column {\tt c} in the FTS table}
\State $w \gets $ weight for column {\tt c}
\State $idf \gets idf + $ $log(\frac{ndoc}{ndocshitcount})\times w $
\State $tf \gets tf + \frac{(nhitcount \times w)}{(nglobalhitcount \times doclen)}$
\EndFor
\EndFor
\State $score \gets $ $\frac{(tf \times idf)}{(k + tf)}$ \\
\Return $score$
\end{algorithmic}

\end{description}
\section{Results}
This section shows results of some of the sample queries on new \textit{apropos}
to demonstrate how it performs as compared to the traditional version.
\endnote{Although this implementation of \textit{apropos} by default returns 10
results for a query but in the following listings the output has been
snipped to save space
} \\

{\tt \small
\lstset{language = XML, caption={Add new user},}
\begin{lstlisting}
$ apropos ``add new user''
ssh-add(1)      adds private key identities to
the authentication agent...on the command line.
If any file requires a passphrase, ssh-add asks
for the passphrase from the user. The
passphrase is read from the user's tty. ssh-add
retries the last passphrase if multiple
identity files are given...

chpass(1)       add or change user database
information
add or change user database information

useradd(8)      add a user to the system
The useradd utility adds a user to the system,
creating and populating a home directory if
necessary. Any skeleton files will be provided
for the new user if they exist in the skel-dir
directory (see the k option). Default...
.
.
.
\end{lstlisting}
}

{\tt \small
\lstset{language = XML, caption={make directory},}
\begin{lstlisting}
$ apropos ``make directory''
make(1) maintain program dependencies
...CURDIR A path to the directory where make
was executed. Refer to the description of PWD
for more details. MAKE The name that make was
executed with argv[0].
For compatibility make also sets .MAKE with the
same value. The...

mkdir(1)        make directories
make directories

ln(1)   make links
...a directory in which to place the link;
otherwise it is placed in the current directory
If only the directory is specified, the link
will be made to the last component of
source_file . Given more than two arguments,
ln makes...

mkfifo(1)       make fifos
make fifos...of a=rw mkfifo requires write
permission in the parent directory. mkfifo
exits 0 if successful, and >0 if an...

mkdir(2)        make a directory file
...will contain the directory has been
exhausted. EDQUOT The user's quota of
inodes on the file system on which the
directory is being created has been
exhausted. EIO An I/O error occurred
while making the directory entry or...
\end{lstlisting}
}

{\tt \small
\lstset{language = XML, caption={signal number to string},}
\begin{lstlisting}
$ apropos ``signal number to string''
psignal(3)      system signal messages
...the signal number is not recognized
sigaction(2) , the string Unknown signal
is produced. The psiginfo function produces
the same output as the psignal function,
only it uses the signal number information
from the si argument. The message strings can...

intro(2)        introduction to system calls
and error numbers
...undefined signal to a signal(3) or kill(2)
function). 23 ENFILE Too many open files in
system . Maximum number...shell. Pathname A
path name is a NUL -terminated character
string starting with an optional slash,
followed by zero or...

dump_lfs(8)     filesystem backup
...low estimates of the number of blocks to
write, the number of tapes it...a string
containing embedded formatting commands for
strftime(3). The total formatted string
is...in. If dump_lfs rdump_lfs receives a
SIGINFO signal (see the status argument of
stty...
\end{lstlisting}
}

\begin{figure*}[htp]
\begin{center}
\includegraphics [scale=0.36]{/home/abhinav/development/AsiaBSDCon/spell.png}
\caption{The spell corrector and the CGI frontend in action}
\label{}
\end{center}
\end{figure*}

In \textbf{Figure 1} the spell corrector can be seen in action along with the web
interface that was developed using \textit{apropos-utils}. \\

\section{Related Work}
There are at least two projects which are related to this in some way.
\begin{description}
\item[man-db] \hfill \\
\textit{man-db} [11] is a complete implementation of the man page documentation
system and it is used on a number of GNU/Linux distributions. \textit{man-db}
takes an interesting approach for indexing the man page data. Unlike the
classical \textit{apropos}, it uses a Berkley DB database but still its index
is limited to the NAME section only. It adds an option `K' to \textit{man(1)} to
allow a crude full text search but it is not very efficient nor effective.
\end{description}

\begin{description}
\item[mandocdb] \hfill \\
\textit{mandocdb} [12] is more in line with the goals of this project but it
takes a novel approach. It indexes the keywords extracted from the man pages in a
key-value store using btree(3) [13]. It also comes with its own implementation
of \textit{apropos} which performs search using this key-value store. The
main key point of this implementation is that it exploits the semantic structure
but rather than indexing the complete content, it plucks out specific pieces of
information from the man pages, for example Author names, function names,
header file names, etc.
from the man pages. It is more sophisticated and much more useful than classic
\textit{apropos}, but it is still essentially a keyword lookup based
implementation.
\end{description}

\section{Future Work}
\begin{description}
\item[Work on Ranking Algorithm] \hfill \\
The ranking algorithm is based on the probabilistic model of information
retrieval [14]results. It
uses certain parameters whose values are usually dependent on the corpus and the
search application. For example, certain weight parameters have been assigned to
different sections of man pages, signifying how important the information in a
particular section is as compared to the rest of the sections. At the moment
these values have been determined
manually but one goal is to use some supervised machine learning techniques in
order to automate this so that more better search results can be obtained.
\end{description}

\begin{description}
\item[Using external storage] \hfill \\
A new feature is under development in \textit{Sqlite}, to support external
storage option in FTS databases. This option would allow building the FTS
index without actually storing any content in the database itself (except for
the data required to maintain the indexes). This will save bulk of the disk
space currently required by \textit{apropos} database. The disk space thus
saved might also allow for experiments with ranking algorithms which require
storing additional metadata on the disk.
\end{description}

\section{Availability}
The code for this project has been imported into the NetBSD source tree and
being maintained there. Experimental work is still being done in the \textit
{github} repository of the project. It can be obtained from:\newline

{\tt
 \emph{https://github.com/abhinav-upadhyay/
 apropos\_replacement}
}

\section{Acknowledgement}
This project has been developed as part of \emph{Google Summer of Code 2011}
[17], so thanks to \textbf{Google} for sponsoring it. Special thanks to
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. We would like to thank \textbf{David Young} who was involved with
this project closely and offered useful help and guidance throughout. A special
thanks goes to \textbf{Thomas Klausner} who helped in writing and reviewing
man pages for this project. Thanks to \textbf{Petra Zeidler} for administering
the GSoC program for \textit{The NetBSD Foundation}. \\

\begin{thebibliography}{20}
\bibitem{1}
\emph{NetBSD manual page for the old version of apropos}
\newline
{\tt
http://netbsd.gw.com/cgi-bin/
man-cgi?apropos++NetBSD-5.1
}

\bibitem{2}
\emph{NetBSD manual page for makewhatis(8)}
\newline
{\tt
http://netbsd.gw.com/cgi-bin/
man-cgi?makewhatis++NetBSD-5.1
}

\bibitem{3}
Brin, S.; Page L.
\emph{The Anatomy of a Large-Scale Hypertextual Web Search Engine}
Computer Networks and ISDN Systems
30:107-117, 1998
\bibitem{4}
Manning; Raghwan; Schutze
\emph{Introduction to information retrieval},
3-9,
2008
\bibitem{5}
Porter, M. F.
\emph{An algorithm for suffix stripping},
Program,
14(3): 130-137,
1980

\bibitem{6}
\emph{mdocml online manual page for
libmandoc}
\newline
{\tt
http://mdocml.bsd.lv/mandoc.3.html
}

\bibitem{7}
\emph{Sqlite home page}
\newline
{\tt
http://sqlite.org
}

\bibitem{8}
\emph{Sqlite FTS3 and FTS4 Extensions}
\newline
{\tt
http://sqlite.org/fts3.html
}

\bibitem{9}
\emph{Online manual page for makemandb(8)}
\newline
{\tt
http://netbsd.gw.com/cgi-bin/
man-cgi?makemandb++NetBSD-current
}

\bibitem{10}
\emph{Online manual page for apropos-utils(3)}
\newline
{\tt
http://netbsd-soc.sf.net/projects/
apropos\_replacement/apropos-utils.html3
}

\bibitem{101}
\emph{man-db, the on-line manual database}
\newline
{\tt
http://man-db.nongnu.org/
}

\bibitem{12}
\emph{Online manual page for mandocdb(8)}
\newline
{\tt
http://mdocml.bsd.lv/mandocdb.8.html
}

\bibitem{13}
\emph{NetBSD online manual page for btree(3)}
\newline
{\tt
http://netbsd.gw.com/cgi-bin/
man-cgi?btree+3+NetBSD-5.1
}

\bibitem{14}
Fuhr, Norbert
\emph{Probabilistic models in information
retrieval}
The Computer Journal
1992
\bibitem{15}
Zaragoza, H.; Craswell, N.; Taylor, M.;
Saria, S.; Robertson, S.
\emph{Microsoft Cambridge at TREC–13: Web
and HARD tracks}
In proceedings of TREC-2004

\bibitem{16}
\emph{Online manual page for man(7)}
\newline
{\tt
http://mdocml.bsd.lv/man.7.html
}

\bibitem{17}
\emph{Google Summer of Code home page}
\newline
{\tt
http://code.google.com/soc/
}

\bibitem{18}
\emph {The mdocml project home page}
\newline
{\tt
http://mdocml.bsd.lv
}

\bibitem{19}
Jones, Sparck K.
\emph {A statistical interpretation of term specificity and its application in retrieval.}
Journal of Documentation
Volume 28 Number 1 1972 pp. 11-21
\newline
{\tt
http://mdocml.bsd.lv
}

\end{thebibliography}

\theendnotes

\end{document}